How has the Brazilian Amazon been constructed as a problem? Presidential speeches and transnational politics since 1985

### 2023-01-26 Submission/Re-submission/draft/draft_resubmission.rmd
Started to update text in rmarkdown

---
title: How has the Brazilian Amazon been constructed as a problem:presidential speeches and transnational politics since 1985
author:
- Livio Silva-Muller
- Henrique Sposito^[Both co-authors contributed equally to the article. Names are ordered alphabetically. Livio is a PhD Candidate in Anthropology and Sociology at the Geneva Graduate Institute and affiliated to the Albert Hirschman Centre on Democracy. Henrique is a PhD Candidate in International Relations and Political Sciences at the Geneva Graduate Institute and affiliated to the Centre for International Environmental Studies.]
date: 23 September 2022
output:
  pdf_document: default
  latex_engine: xelatex
  fig_caption: yes
  html_document:
    df_print: paged
fontsize: 12pt
mainfont: Times New Roman
header-includes:
   - \usepackage{floatrow}
   - \floatsetup[figure]{capposition=top}
   - \usepackage{float}
   - \floatplacement{figure}{H}
   - \usepackage[T1]{fontenc}
   - \usepackage[utf8]{inputenc}
   - |
    \providecommand{\keywords}[1]{\textbf{Keywords:} #1}
    \providecommand{\wordcount}[1]{\textbf{Word Count:}#1}
always_allow_html: TRUE
bibliography: references.bib
csl: chicago.csl
abstract: Presidential speeches can influence the ways people think and act towards the environment , but they are understudied. We propose a framework to investigate how the Brazilian Amazon has been constructed as a problem across time and space. Using supervised machine learning, we classify statements about the Amazon in 6240 transnational presidential speeches since 1985. We find that national and international events drive the frequency at which the topic of the Amazon is mentioned in presidential speeches. While constructing the Amazon as a problem of economic integration are the most common until the mid-2000s, environmental conservation and social development constructions temporarily surpass economic integration from 2010 to 2015. In turn, constructing the Amazon as an issue of sovereignty increases since 2010. Lastly, presidents usually construct the Amazon as an issue of environmental conservation internationally. When speaking within the Amazonian region, presidents tend to construct the Amazon as an issue of social development and economic integration.
---

@@

\pagebreak
# 1. Intro

# 2. Theory

# 3. Methods

## 3.1 Problem-construction

@@

@@

@@

@@

models <- tibble::tibble("Models Description" = c("Fixed-effects logistic regressions indexed by president (controlling for time and ideology)."),
                         "Dependent variables" = c("Pure environmental conservation; pure economic integration; pure social development; and pure national sovereignty."),
                         "Independent Variables" = c("Amazonian states within Brazil (reference category); non-Amazonian Brazilian states; Brasilia; Amazonian countries; and international."),
                         "Controls" = c("Yearly deforestation rates lagged by one year; average annual inflation; election years; and peak years for Amazon appearing year in presidential speeches (footnote)."))

kbl(models, booktabs = T, caption = "Model description") %>%
  kable_styling(latex_options = c("striped", "repeat_header",
                                  "scale_down", "HOLD_position"),
                position = "center", font_size = 9, full_width = TRUE)
```

@@

```{r Figure pure, fig.cap="Pure-types in time", echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE, include=TRUE}
mmt1 <- ama_mx %>%
  filter(mx_cat2 == "Mixed-types") %>%
  count(mx_cat) %>%
  mutate(perc = n / sum(n)) # averages in relation to only mixed-types
mmt2 <- ama_mx %>%
  count(mx_cat)%>%
  mutate(perc = n / sum(n)) # averages in relation to full sample
ama_mx %>%
  count(mx_cat, mx_cat2, year) %>%
  group_by(year) %>%
  mutate(perc = n / sum(n)) %>% 
  filter(mx_cat2 == "Pure-types") %>%
  ggplot() +
  geom_smooth(aes(x=as.numeric(year), y = perc, color = mx_cat, fill = mx_cat),
              se = FALSE, linewidth =0.75, linetype="solid") +
  scale_y_continuous(labels = percent_format(accuracy = 1)) +
  labs(x = "", y = "", title ="Share of each pure-type problem-construction by year",
       subtitle = "Sample composed by 1924 Amazonian statements in presidential speeches since 1985",
       caption = "Curves in the plot were smoothed using loess method.") +
  theme_am() +
  # theme(legend.position = "bottom")+
  # guides(fill=guide_legend(nrow=2,byrow=TRUE)) +
  facet_wrap("mx_cat") +
  geom_smooth(amazon_def_year, mapping = aes(x = year, y = def_year/100),
              color= "darkolivegreen", linetype="twodash", se = FALSE) +
  annotate("text", x = 1989, y = 0.16, label = "Deforastation", size = 2) +
  annotate("text", x = 2001, y = 0.18, label = "27772 km2\ndeforested" , size = 2)+
  annotate("text", x = 2014, y = 0.03, label = "4571 km2\ndeforested", size = 2)
```

>>>>

---
title: How has the Brazilian Amazon been constructed as a problem? Presidential speeches and transnational politics since 1985
author:
- Livio Silva-Muller
- Henrique Sposito^[Both co-authors contributed equally to the article. Names are ordered alphabetically. Livio is a PhD Candidate in Anthropology and Sociology at the Geneva Graduate Institute and affiliated to the Albert Hirschman Centre on Democracy. Henrique is a PhD Candidate in International Relations and Political Sciences at the Geneva Graduate Institute and affiliated to the Centre for International Environmental Studies.]
date: 27 January 2023
output:
  pdf_document: default
  latex_engine: xelatex
  fig_caption: yes
  html_document:
    df_print: paged
fontsize: 12pt
mainfont: Times New Roman
header-includes:
   - \usepackage{floatrow}
   - \floatsetup[figure]{capposition=top}
   - \usepackage{float}
   - \floatplacement{figure}{H}
   - \usepackage[T1]{fontenc}
   - \usepackage[utf8]{inputenc}
   - |
    \providecommand{\keywords}[1]{\textbf{Keywords:} #1}
    \providecommand{\wordcount}[1]{\textbf{Word Count:}#1}
always_allow_html: TRUE
bibliography: references.bib
csl: chicago.csl
abstract: Presidential speeches can influence the ways people think and act towards the environment, but they are understudied. We propose a framework to investigate how the Brazilian Amazon has been constructed as a problem across transationally across settings. Using supervised machine learning, we classify statements about the Amazon in 6240 transnational presidential speeches since 1985. We find that national and international events drive the frequency at which the topic of the Amazon is mentioned in presidential speeches. While constructing the Amazon as a problem of economic integration are the most common until the mid-2000s, environmental conservation and social development constructions temporarily surpass economic integration from 2010 to 2015. In turn, constructing the Amazon as an issue of sovereignty increases since 2010. Lastly, presidents usually construct the Amazon as an issue of environmental conservation internationally. When speaking within the Amazonian region, presidents tend to construct the Amazon as an issue of social development and economic integration.
---

@@

\pagebreak

# 1. Introduction

*We need to protect the Amazon from foreign interests. We need to exploit the Amazon's natural resources. We need to provide better living standards for the people in the Amazon. We need to preserve the Amazon as a standing ecosystem*. Each of these statements contains an implicit assumption of what needs to be solved, or in other words, it constructs the Amazon as a particular problem. Scholars studying the Brazilian Amazon often identify different governments throughout Brazilian history as proponents of a view of the region, its people, and forest [@drummond2006; @padua2012; @franchini2019; @capobianco2019; @pereira2021]. While the military dictatorship (1964-85) is, for example, associated with understanding the Amazon as issues of national sovereignty and economic integration, Sarney’s (1985-1990) and Lula’s (2003-2010) presidencies are often tied to environmental conservation. Yet, such monolithic representations advance the view that governments understood and acted consistently towards the Amazon as a specific problem, masking government diversity of discursive constructions and policies.

In this article, we show how the Brazilian Amazon has been constructed as a problem transnationally in presidential discourses since 1985. Building on @bacchi2009's theory of problem-representation and @keck1998's theory of transnationalism, we offer a framework that ties discursive problem issue representations to relationships that transcend nation-states to identify and explain variation in problems-constructions. We shift the focus from governmental policy to discourse to capture a broader range of understandings for policy objects, problems that demand dedicated policy attention, that introduce, shape, and justify outcomes to broad audiences [@zarefsky2004; @gillion2016]. Although problem-constructions take place in a series of instances, we analyze the case of official presidential discourses. The Brazilian executive branch of government, led by the president, has historically been the primary actor in shaping public policy in Brazil [@pinto2016]. Working with presidential discourses also allows us to identify variations in how the Amazon is discursively constructed in transnational politics beyond environmental policies. Whereas understudied in the Amazonian context, presidential speeches legitimize perceptions about the Amazon that are key for policy adoption and implementation [see @alesina2009; @lopez2020]. This is especially pertinent for deforestation in the Amazon as expectations of local actors, generated from material and discursive governmental practices, are crucial in decisions to deforest on the ground [@assuncao2015; @capobianco2019; @capobianco2021; @campbell2015]. As crucial participants in transnational politics, presidents deploy different problem-constructions according to the transnational audiences they are speaking to.

To investigate how the Brazilian Amazon has been constructed as a problem in presidential speeches, we create a dataset containing 6240 official presidential speeches by all Brazilian presidents since 1985. Next, we identify all instances where presidents refer to the Amazon in their speeches. We develop a codebook grounded in Amazonian historiography to code how each of these statements constructs Amazon as a particular problem. With this codebook, we manually code a randomly selected training set of statements and use supervised machine-learning to label the remaining observations. We then conduct a descriptive analysis of when, where, and how presidents construct the Amazon as a problem.

Our findings are threefold. First, we find that the frequency in which the Amazon appears as a topic in presidential speeches is driven by domestic and international environmentally-related  events. Second, we find that presidents are more likely to construct the Amazon as an issue of environmental conservation outside of the Amazon region, this is especially true when presidents speak internationally. Within the Amazon region, presidents tend to construct the Amazon as an issue of social development and economic integration. Third, the same government adopts a diverse range of problem-construction. While the constructions of the Amazon as an issue of economic integration were more salient from 1985 to the mid-2000s, environmental conservation and social development problem-constructions steadily grew since the 2000s, temporarily surpassing economic integration as a more salient construction from 2010 to 2015. Constructing the Amazon as an issue of sovereignty steadily increased after 2010. This article provides the first comprehensive overview of when, where, and how the Amazon is constructed as problems in presidential speeches.

# 2. Theory: problem-constructions, transnational settings, and presidential speeches

## 2.1 Governments and problem-constructions

Social scientists often describe federal governments as proponents of a cohesive set of policies toward the Amazon [@drummond2006; @padua2012; @franchini2019; @pereira2021]. The 1964 military dictatorship, for example, is associated with securing sovereignty in the region by populating it and integrating it into the national economy [@drummond2006]. Whereas governments from the late-1980s up to the late-2000s are associated with a turn towards policies that focus on environmental conservation [@padua2012]. The presidencies of Dilma (2011-2016) and Temer (2016-2018) are connected to the de-prioritizing of environmental conservation policies, and Bolsonaro (2019-2022) is generally associated to the dismantling of environmental policies [@capobianco2019]. Although these works are important to understand how different governments acted towards the Amazon, they represent governments monolithically by bundling them into policy periods, or policy cycles, with one all-encompassing view of the Amazon.

We know, however, that there is more variation within a specific government policy, especially if we consider the Amazon a policy object demanding dedicated policy attention. For instance, while many argue that environmental activism in Brazil emerged despite the military dictatorship [@hochstetler2000; @wood1993], @rajao2021 demonstrates how the strengthening of the Forest Code in 1964 was also a military priority. Furthermore, a substantial part of Lula’s administration (2003-2010) effectiveness in reducing deforestation in the Amazon was achieved because of deforestation leakage to other biomes (Garret, 2022). At the same time, Lula’s administration (2003-2010) provided record high subsidies to the expansion of agriculture in Brazil [@capobianco2021]. That is, we must look at the Amazon beyond the environmental domain as a forest, region, and peoples.

We argue that the reason behind these monolithic representations of governments is conceptual: many studies of the Brazilian Amazon contain an explicit, or implicit assumption, of what the Amazon problem is. @bacchi2009[p. 10] posits that policies have a cultural dimension as they take “shape within specific historical and national or international contexts”. The existence or proposal of a policy implies that there is a (public) problem that needs (governmental) action to be fixed. The alleged problem is not always explicitly stated in a policy, but policies are represented as solutions for both explicit and implicit problems. If the Amazon is represented as an “environmental” problem, as is often the case in the literature, then it demands an “environmental” solution. Consequently, studies focus on understanding environmental policies [@capobianco2019], the work of environmental NGOs [@silva-muller2022], or the rise of global environmental governance [@andonova2022]. Alternatively, studies looking into the developments of social and fiscal policy in the region during the last decades are scarce [@acker2017]. We acknowledge the quality and importance of all these studies. Still, we argue that it is crucial for scholars to study the Amazon as a biological, social, political, and geographical policy object. We conceptualize the Amazon as a policy object that demands dedicated policy attention of a specific nature. Hence, governments can highlight diverse explicit and implicit problems related to a policy object that need political attention. This is what we label as problem-constructions.
<!-- Are we missing a transition from problem-constructions generally to the case of speeeches? -->

## 2.2 Transnational settings and the variation in problem-constructions

The possibility of different problem-constructions for the same policy object in political discourses, requires an explanation of why they would vary. On the one hand, @putnam1988 theory of two-level games in international negotiations can be adapted to explain variation in presidential speeches. @putnam1988 argues that the outcomes of international negotiations lie within the overlap between the interests (and pressure) of domestic and international groups. At the national level, interest groups pressure the government to adopt favorable policies. Internationally, the government seeks to fulfill domestic preferences “while minimizing the adverse consequences of foreign developments” [@putnam1988, p. 434]. While the theory has been expanded to cover multi-levels games [see @collinson1999], its core remains the same: the outcome of negotiations should lie within the overlap of diverse interest groups.

This two-level approach is, for example, the avenue pursued by @calderwood2019 and @calderwood2020 who investigates variation in American presidential speeches about climate change. Consistent with the two-level games, the author finds that presidents are more likely to mention climate change outside the US, though the location of the speech (domestic or foreign) has a small effect on the content of the mention [@calderwood2020]. However, the party affiliation of American presidents mediate how they speak about climate change [@calderwood2019]. Thus, adopting a two-level approach to analyze diverse discursive constructions implies that whatever presidents say inside and outside, lies within the policy expectations of domestic and foreign interest groups. Consequently, presidential speeches are not necessarily contradictory across venues but a common denominator of the interest of different groups.

On the other hand, @keck1998 theory of international politics as transnational networks, developed in the context of value-based advocacy groups, provides an alternative explanation for variation in problem-constructions. The authors reject the dichotomy of international and domestic levels arguing that the identity and goals of state and non-state actors are not derived from their structural location vis-a-vis domestic and foreign interest groups. Rather, international politics as transnational networks have “a structured and structuring dimension” [@keck1998, p. 4]: state and non-state actors participate in and shape international politics. In this sense, the positions state actors as components of a transnational network derive from complex interactions in which their identities and goals that are malleable and can vary across time and space.

Important participants of transnational networks, as presidents, have a myriad of tactics to introduce, define, persuade, socialize, and pressure when interacting with other actors. Presidents can, for example, engage in information politics by creating politically meaningful information and/or moving it where they believe it can have an impact [@keck1998, p. 24]. As well, presidents can engage in leverage politics by leveraging institutions to realize symbolic or material gains. Transnational networks do not assume a consistent common denominator of problem-constructions. That is, presidents construct problems and attempt to inform or leverage transnational audiences for symbolic or material gains according to their expectations of said audience. This indicates that how presidents construct policy objects as problems can vary within the country (i.e. speaking at diverse states or different associations) as well as when speaking internationally (i.e. during a climate conference or after a catastrophe). We call these levels in which there could be variation in how presidents construct a problem a setting. The setting is a level of aggregation that encompasses one, or multiple, audiences and makes sense concerning the policy object under study, in this case, the Amazon.

## 2.3 Problem-constructions and policy

In addition to the setting where the speech takes place, it is natural that outcomes in policy adoption and implementation can affect when and how presidents construct problems. @gillion2016 argues that there are two types of outcomes that can affect presidential discourses: institutional and societal. An institutional outcome relates to how presidential discourses can shape policy coalitions and the policy agenda. While the relationship between presidential speeches and policy may not be causal, they have the power to introduce, define, and justify public policy, as well as shape issue perception to broad audiences [@zarefsky2004]. This is particularly true in the Brazilian context where the executive branch of government, led by the president, have historically been the primary players in shaping public policy, especially after the 1988 constitution [@pinto2016].

A societal outcome, instead, relates to how discourse informs attitudes and behavior of people on the ground and affects policy outcomes. In the case of deforestation, for example, Amazonian scholars argue that the expectations of local actors generated from governments’ decisions [@assuncao2015] and discourses [@capobianco2019; @capobianco2021; @campbell2015] are a crucial factor in choice to deforest on the ground. As presidents participate and shape transnational politics by constructing the Amazon as a certain issue, discourses define policy agendas, direct “political officials’ at multiple stages of the public policy-making process” [@gillion2016, 19], influence people’s decision on the ground, and justify previous decision or policy outcomes.

Considering the possibility of varied problem-construction, the relationship between policy and discourse becomes endogenous and co-constituted: problem constructions in presidential speeches can justify past outcomes and/or lead to future outcomes. A bad track record in climate change policy outcomes can be justified with a good track record in energy security. This was, for example, the strategy adopted by President Bush during the Kyoto Protocol negotiations [@calderwood2020]. As presidents participate in transnational networks, with malleable identities and goals, will refer to outcomes interchangeably, based on their expectations about the setting they are speaking at and the transnational audiences present. Policy outcomes, therefore, can be (re)constructed in presidential discourse in diverse forms to shape certain problem definitions and possible solutions.

# 3. Research Design: operationalizing Amazonian problem-constructions

## 3.1 Codebook (Amazonian historiography)

To capture the cultural dimension and historical contexts behind problem constructions, we rely on the Amazonian historiography literature. We refer to Amazonian historiography as the body of research conducted by social and environmental scientists that tell the story of diverse governmental programs that studies the Amazon region, forest, and its people. Within this literature, we identify three historical-political constructions: national sovereignty, economic integration, and environmental conservation. We argue a fourth separate one also exists: social development. The social development construction focuses on the constitutional rights and dignity of the inhabitants of the Amazonian region.

### National Sovereignty

In the process of securing Amazonian borders since the 18th century, Brazil thwarted “the imperial ambitions of France, Britain, the United States, Belgium, Bolivia, and Peru” [@hecht1990, p. 8], and when the dust settled and the scramble was over, more than half of the Amazon territory emerged Brazilian. While the Brazilian military diplomacy following the scramble was very successful, the process did not come without its traumas as, for example, the negotiations with Bolivia in 1902 to secure the state of Acre [@hecht1990]. These traumas were referred to, and offered context, to Amazonian discourses and policies that highlighted Brazil’s sovereignty over the Amazon [@hecht2020]. More recently, as we move to a world where non-state actors gain importance in international politics [@silva-muller2022; @andonova2014; @keck1998], sovereignty-related problems become more varied. Threats to national sovereignty, consequently, broaden from nation-states to a wide set of actors. The sovereignty problem-construction advances the view that the Brazilian Amazon belongs to Brazil and any foreign or non-state presence in the region can be part of a broad strategy to take the region. The policy solutions to the issue of sovereignty included the monitoring of the borders and strict regimes related to entry into the region.

### Economic Integration

The Vargas dictatorship (1937-46) and the military dictatorship (1964-89) took over the task of modernizing the Amazon [@capobianco2019; @becker2005]. In 1966, the Brazilian Military launched ‘Operation Amazon’, a policy to modernize the region based on three assumptions [@acker2021]. First, it assumed that nature should be conquered by mankind. Second, it assumed that exploiting natural resources would render the Amazon region economically profitable. Third, it assumed that populating the region was necessary to integrate it into the country and exert control over the territory. Concretely, this meant a series of major infrastructure projects, incentives for settlers to expand the agricultural frontier, and the establishment of a tax-free zone in Manaus to attract industry. The economic integration problem-construction advances the view that the Brazilian Amazon needs to be developed and modernized. These policy solutions have at their core the development of the necessary infrastructure (physical, fiscal, or monetary) to integrate the region into the national and international economy.

### Environmental Conservation

The rapid economic changes in the region from the 1960s to the 1980s were matched with the birth of environmental institutions [@drummond2006]. Arguably, the creation of these institutions stemmed from the impression of the lack of control over the environment following the efforts to economically integrate in the region [@acker2021; @capobianco2021; @hecht1990]. This process accelerated in the late 1980s with the birth of modern environmentalism [@viola1987], epitomized in the 1992 Earth Summit in Rio de Janeiro [@hochstetler2021; @capobianco2021]. The environmental conservation problem-construction emphasizes that the Amazon should be preserved, deforestation should be halted, and the sustainable practices of indigenous and local peoples should be maintained through the protection of their territories and rights to self-determination [@hochstetler2007]. The policy solution implies, for example, further investments in command-and-control institutions, the valuation of standing ecosystems through incentive schemes, and the creation of protected areas.

### Social Development

Governments can emphasize the lack of hospitals, sanitation, and schools concerning peoples’ dignity, standards of living, and other constitutional rights. However, such emphasis can be masked within both the economic integration and environmental conservation accounts. While for some groups, such as indigenous peoples, constitutional rights and preservation of the forest go hand in hand, providing better standards of living to large portions of populations of the Amazonian states and conservation are not necessarily connected. As democracy established itself in Brazil, especially with the 1988 constitution, settlers in the Amazon became electorates with constitutional rights. Social development is a problem-construction that focuses on constitutional rights, citizenship, and dignity. The policy solution to social development implies the policies to facilitate access to water, sanitation, electricity, internet, and radio, as well as the construction of schools and hospitals locally.

@@

To identify varied problem-constructions, we focus on presidential speeches. By focusing on presidential speeches, rather than specific policy domains, we analytically avoid hinging our analysis on an underlying assumption of what the problem is. Thus, we are not necessarily prioritizing an environmental, economic, or sovereignty-based view of the Amazon. We build upon @cezar2020 dataset, which contains all official speeches by Brazilian Presidents from 1985 to 2019. We update the dataset by scraping and adding all official speeches from 2020 to 2022. The final dataset encompasses 6240 speeches. We proceed to identify speeches that refer to the Amazon as a region, people, or forest. We do so by detecting all speeches in which the stem “amazon” appears. In Portuguese, the stem captures terms such as “Amazonia”, “Amazonica”, “Amazonidas”, “Amazonense(s)”, “Amazonas”, among others. We find that 968 speeches are, at least partially, about the Amazon. Using the poldis R package [@sposito2021], we extract two sentences before and after the sentence in which the stem “amazon” appears. We opt for picking two sentences around, rather than words, because sentences usually contain a cohesive idea. This is how our unit of analysis, Amazonian statements, are generated. This process yields 2048 unique Amazonian statements about the Amazon. On average, an Amazonian statement contains 131 words.

We code Amazonian statements in three steps. First, we develop a codebook for coding Amazonian problem-constructions identified with the Amazonian historiography (see appendix). In their conceptualization, each problem-construction is mutually exclusive; that is, they cover different forms of constructing the Amazon as a problem. However, an Amazonian statement can be coded as referring to multiple problem constructions (see Figure 1 below). Amazonian statements, thus, can be pure-type statements (i.e. constructs the Amazon as a single problem) or mixed-types (i.e. constructs the Amazon as multiple issues).

@@

Second, with the codebook in hand, each one of the authors separately hand-coded 1024 randomly selected Amazonian statements. This amount refers to 50% of all the Amazonian statements identified. We chose to hand code half of the observations because there are several nuances in how presidents talk about the Amazon over time. As well, since we have four different code categories, the size of the training set should be increased [see @grimmer2022]. The hand-coding strategy provides a robust and reliable number of observations to be used for training and validation of the automated model. Once hand coding was completed, the average intercoder agreement for hand-coded categories was 85%. For each non-matching coded observation, the authors discussed and sorted their disagreements, which were subsequently specified in the codebook.

Third, the hand-coded data is randomly divided into a training set, containing 80% of the data (819 observations), and a validation set, containing the remaining 20% of the data (205 observations). We employ a support-vector machine (SVM) algorithm, a non-probabilistic linear classifier appropriate for binary categories, to label texts [@meyer2021; @noble2006]. To validate the model, the SVM algorithm is trained using the hand-coded training set and employed to classify observations in the validation set. As a final robustness check, the authors randomly select 100 automatically coded statements and find that the automated model is 95% accurate in coding these statements as the authors would. Automating the coding of the remaining half of the observations is estimated to have saved the authors over one month of work in comparison to manual coding. The final dataset for analysis, excluding false positive matches, contains 1924 coded Amazonian statements.

@@

To analyze our data, we model the share of a specific pure problem-construction, as the dependent variable, relative to the settings in which a particular speech happens, as our independent variable. This means we predict the probability of a president constructing the Amazon as an issue of (1) environmental conservation, (2) economic integration, (3) social development, and (4) national sovereignty, separately, and in relation to the remaining problem-constructions (see table 1 below). We focus on single problem constructions to provide unambiguous interpretations and more robust findings. We divide settings in terms of non-Amazonian Brazilian states, Amazonian states within Brazil, Brasilia, Amazonian countries, and international. This level of aggregation refers to geo-political settings that make sense in light of the Amazon, as our policy object. This aggregation combines different expectations presidents can have about transnational audiences in each of these settings. Brasilia, specifically, is treated as a separate setting due to its unique location and since most presidential speeches are delivered there. The model controls for yearly deforestation rates, inflation, election years, and peak years for Amazon in presidential speeches.

@@

## Count average words in Amazonian Statements
# amazon_speeches_long <- readRDS("~/Documents/GitHub/amazondef/Replication_Files/amazon_speeches_long.Rds")
# sum(stringr::str_count(amazon_speeches_long$AM2, '\\w+'))/2048
## 131 on average

## Comparison with inflation, criminality, unemployment, and inequality
# pop %>%
#   mutate(desigual = ifelse(grepl("desigual", text,
#                                  ignore.case = TRUE), 1, 0),
#          crime = ifelse(grepl("violencia|crime|criminalidade|seguranca publica",
#                               text, ignore.case = TRUE), 1, 0),
#          inf = ifelse(grepl("inflacao|inflaçao",
#                              text, ignore.case = TRUE), 1, 0),
#          unemploy = ifelse(grepl("desempreg",
#                              text, ignore.case = TRUE), 1, 0)) %>%
#   summarise(desigual = mean(desigual),
#             crime = mean(crime),
#             inf = mean(inf),
#             unemploy = mean(unemploy))

models <- tibble::tribble(~"", ~Description,
                          "Model", "",
                          "Fixed-effects logistic regressions indexed by the president", "The regression method allows to control for variables not measured (e.g. ideology) related to individuals (i.e. presidents) [see @allison2009].",
                          "Dependent variables", "",
                          "Problem constructions", "Pure environmental conservation; pure economic integration; pure social development; and pure national sovereignty.",
                          "Independent variable", "",
                          "Settings", "Setting where speech occurs as (1) non-Amazonian Brazilian states, (2) Brasilia, (3) Amazonian countries, and (4) international, in relation to the reference category of Amazonian states within Brazil.",
                          "Control variables", "",
                          "Deforestation rates", "Lagged yearly deforestation rates as a control for the possible relationship between deforestation rates and Amazonian problem-constructions.",
                          "Election year", "Election years are controls for possible discursive changes related to the Amazon for these years.",
                          "Inflation rate",  "Average yearly inflation rates as a control for a possible relationship between economic conditions/fluctuations and problem constructions.",
                          "High-profile environmental event", "Years in which high-profile environmental events occurred control for the frequency and, possibly, how the Amazon was constructed as a problem in discourse for those years.")

kbl(models, booktabs = T, caption = "Model description") %>%
kable_styling(latex_options = c("striped", "repeat_header", "scale_down"),position= "center", font_size = 9, full_width = TRUE) %>% row_spec(c(1,3,5,7), bold = TRUE, italic = TRUE)
# see also: pack_rows("Model", 1, 2)
# https://cran.r-project.org/web/packages/kableExtra/vignettes/awesome_table_in_pdf.pdf
```

The main limitations of our approach are threefold. First, related to our subsetting and coding strategies, we classify statements as Amazonian based on a dictionary composed of a single lexicon stem: “amazon”. We chose to do so knowing that a few speeches about the Amazon might not contain the lexicon “amazon”, for example, when the president says, “the forest” or “deforestation”. However, we consider this safer as we cannot be sure that mentions of the forest or deforestation do not correspond to other biomes such as the Cerrado or the Atlantic Forest.

Second, our dataset covers only what is considered an official speech. Presidents, though, give interviews, appear in debates, talk at campaign rallies, and, more recently, post on social media. Problem-construction within presidential discourse, thus, happens in different public, private, and online settings for which we do not account or differentiate in this paper. Furthermore, we aggregate levels to socio-political-cultural settings that make sense in light of the policy object in question and the data at hand. Therefore, we are aware that a category as “international”, for instance, encompasses speeches delivered by presidents in a wide variety of countries, institutions, and venues to diverse audiences, which are not differentiated.

Third, as presidents participate and shape transnational politics, we assume that presidential discourse, policies, and outcomes are co-constitutive. Discourses capture general problem constructions that can be used to introduce, discuss, and/or justify policy and outcomes. Throughout the analysis we carefully discuss how presidential discourses relate to, for instance, deforestation rates and political outcomes by pointing at some of the patterns and correlations but without making any causal claims.

@@

```{r Figure pure, fig.cap="Pure-types in time", echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE, include=TRUE}
mmt1 <- ama_mx %>%
  filter(mx_cat2 == "Mixed-types") %>%
  count(mx_cat) %>%
  mutate(perc = n / sum(n)) # averages in relation to only mixed-types
mmt2 <- ama_mx %>%
  count(mx_cat)%>%
  mutate(perc = n / sum(n)) # averages in relation to full sample
ama_mx %>%
  count(mx_cat, mx_cat2, year) %>%
  group_by(year) %>%
  mutate(perc = n / sum(n)) %>% 
  filter(mx_cat2 == "Pure-types") %>%
  ggplot() +
  geom_smooth(aes(x=as.numeric(year), y = perc, color = mx_cat, fill = mx_cat),
              se = FALSE, linewidth =0.75, linetype="solid") +
  scale_y_continuous(labels = percent_format(accuracy = 1)) +
  labs(x = "", y = "", title ="Share of each pure-type problem-construction by year",
       subtitle = "Sample composed by 1924 Amazonian statements in presidential speeches since 1985",
       caption = "Curves in the plot were smoothed using loess method.") +
  theme_am() +
  theme(legend.position = "bottom")+
  guides(fill=guide_legend(nrow=2,byrow=TRUE))
```

```{r Figure puretypetime, fig.cap="Pure-types across setting in time", echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE, include=TRUE}
amazon_def_year <- amazon_def_year%>%
  mutate(extremes= ifelse(year == 2004|year == 2012, "yes","no"))
ama_mx %>%
  count(mx_cat, location_cat, year) %>%
  group_by(year) %>%
  mutate(perc = n / sum(n)) %>% 
  filter(location_cat == "Amazonian States"|
           location_cat == "International",
         mx_cat == "Pure Economic Integration"|
           mx_cat == "Pure Environmental Conservation"|
           mx_cat == "Pure Social Development") %>%
  ggplot() +
  geom_smooth(aes(x=as.numeric(year), y = perc, fill=),
              se = FALSE, linewidth =0.75, linetype="solid", color="black") +
  scale_y_continuous(labels = percent_format(accuracy = 1))+
  facet_wrap(~location_cat+mx_cat, nrow=2)+
  labs(x = "", y = "", title ="Relationship between problem-constructions and deforestation rates",
       subtitle = "Bars indicated yearly deforestation level and curves indicated yearly share of problem-construction.",
       caption = "Curves in the plot were smoothed using loess method.")+
  theme_am()+
  geom_col(amazon_def_year, mapping = aes(x = year, y = def_year/100,fill=extremes),alpha=.5)+
  scale_fill_manual(values=c("lightgrey", "darkolivegreen"))+
  annotate("text", x = 2008, y = 0.3, label = "27'772 km2\ndeforested" , size = 3, color="darkolivegreen", face="bold")+
  annotate("text", x = 2016, y = 0.04, label = "4'571 km2\ndeforested", size = 3, color="darkolivegreen", face="bold")
```
